---
title: "Pokemon Status and Type"
subtitle: "Chi-Square Test of Association between variables Pokemon Status and Type"
author: "Jeremy Jian Zhi Chia (s3917515)"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output:
  slidy_presentation:
    font_adjustment: 0
    highlight: haddock
    mathjax: "default"
    df_print: paged
---

```{r setup, include=FALSE}
# Use this chunk to quietly load your pacakges

knitr::opts_chunk$set(
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)
# Packages loaded in this chunk will not appear in the presentation. 

library(ggplot2) # Useful for creating plots
library(dplyr)  # Useful for data maipulation
library(knitr) # Useful for creating nice tables

```


# RPubs link information

- You must publish your presentation to RPubs (see [here](https://astral-theory-157510.appspot.com/secured/RBootcamp_Course_04.html#creating_an_r_markdown_document_in_r_studio)) and add this link to your presentation here.

- Rpubs link comes here: www.........

- This online version of the presentation will be used for marking. Failure to add your link will delay your feedback and risk late penalties.

# Introduction

- The following link will help you with creating R Markdown Slidy Presentations
- http://rmarkdown.rstudio.com/slidy_presentation_format.html
- Don't forget about [Bootcamp 4](https://astral-theory-157510.appspot.com/secured/RBootcamp_Course_04.html)
- A good introduction provides a brief background to the problem, defines important terms, and leads to a strong rationale.

- The purpose of this investigation is to investigate if there are any association between Pokemon Status and Pokemon Type.
- Pokemon Status gives the status of any given Pokemon, namely Normal, Mythical, Sub-Legendary or Legendary
- Each Pokemon has a type which determines its weaknesses/resistance to attacks. Some example types include Grass, Fire, Water, Psychic, Electric etc.


# Introduction Cont.

- Keep everything short and straight to the point. 
- Use bullet points to help minimise text.
- Add relevant images to make your presentation more appealing
- Remember, you have a maximum of 20 slides to fit everything in.
- Ensure each slide fits on one screen. The reader should not have to scroll down. 

- As I grew up watching Pokemon, playing Pokemon video games and collecting Pokemon cards, I thought it would be fun and reminiscent to choose this dataset and be able to apply statistical tests on it to gain some insights!


# Problem Statement

- State the overall problem/question driving the investigation
- Summarise how you will use statistics to solve the problem or answer your question.

- As all Pokemon are associated with a status and type, I'd like to find out if there exists a relationship between the Status and Type.
- Will knowing the Pokemon Type allow us to predict the Pokemon Status, or will knowing the Pokemon Status allow us to predict its Type?
- Does any Pokemon Status make up of predominantly a certain Pokemon Type?
- Does any particular Pokemon Type have a significant number of Legendary Pokemons?
- The Chi-square Test of Association is used to test if there is a relationship between the two Status and Type.
- This test will allow us to know if the Status and Type are independent/dependent on each other and whether knowing the values of Pokemon Status helps to predict the value of Pokemon Type and vice versa.

# Data

- You can use multiple slides to break up your sections. 
- Explain how you collected your data.
- There should be enough detail here so that someone else could replicate your data collection.
- Explain the sampling method if known
- Ensure you reference the data source if you have used Open Data

- Complete Pokemon Dataset (All 8 generations)
- Source: https://www.kaggle.com/mariotormo/complete-pokemon-dataset-updated-090420
- This dataset contains records of Pokemons across all eight generations.


# Data Cont.

- List and explain the important variables. 
- If you have factors, explain each level.
- Explain the scale of numeric variables.
- Justify everything that you do to preprocess the data.

- This dataset contains rich information on each Pokemon, they include names in various languages, primary and secondary type, height and weight, special abilities, various battle metrics (Attack, Defense, Speed etc.), egg type, gender distribution and various attack damage against any Pokemon Type.
- Variables of interest: Status and Type
- The structure is inspected and Status and Type are set as unordered factors
- Unique values of variables of interest are inspected to ensure there are no errors
- Existence of null values are also checked

```{r}
# Check for NAs
sum(is.na(pokedex_Update_05_20_$status))
sum(is.na(pokedex_Update_05_20_$type_1))
```
```{r}
pokedex_Update_05_20_$type_1 <- pokedex_Update_05_20_$type_1 %>% as.factor()
pokedex_Update_05_20_ %>% str()
```
# Descriptive Statistics and Visualisation

- Summarise the important variables in your investigation.
- Use visualisation to highlight interesting features of the data and tell the overall story.
- Explain how you dealt with data issues (if any), e.g. missing data and outliers.
- Here are the examples of R chunks and outputs


- You can use the `knitr:kable` function to print nice HTML tables. Here is an example R code:

```{r}
table1 <- table(pokedex_Update_05_20_$type_1, pokedex_Update_05_20_$status)
knitr::kable(table1)
```
```{r}
table2 <- table(pokedex_Update_05_20_$type_1, pokedex_Update_05_20_$status) %>% prop.table(margin = 2)
knitr::kable(table2)
```



# Decsriptive Statistics Cont.

```{r}
ggplot(pokedex_Update_05_20_) +
  aes(x = type_1, fill = status) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(title = "Proportion of Pokemon Status for each Pokemon Type", x = "Type", y = "Proportion")
```
The above barplot displays the proportion of Pokemon status for each Pokemon Type.
- The Normal Pokemons make up the biggest proportion of all Pokemon Types.
- Most Pokemon types have a small proportion of Legendary Pokemons, with the Dragon type having the largest proportion.
- Most Pokemon types have a small proportion of Mythical Pokemons, with the Steel and Psychic type making the largest proportions.
- Most Pokemon types have a small proportion of Sub Legendary Pokemons, with the Flying type having the largest proportion.


# Hypothesis Testing

- Apply an appropriate hypothesis test for your investigation. 
- Ensure you state and check any assumptions. 
- Report the appropriate values and interpret the results.
- Here is an example:

```{r}
model1 <- lm(Sepal.Length ~ Petal.Length, data = iris)
model1 %>% summary()
```

# Hypthesis Testing Cont.

- Here are the examples of mathematical equations:

$$H_0: \mu_1 = \mu_2 $$

$$H_A: \mu_1 \ne \mu_2$$

$$S = \sum^n_{i = 1}d^2_i$$

- You can also place equations inline: $z = \frac{x - \mu}{\sigma}$

# Discussion

- Discuss the major findings of your investigation
- Discuss any strengths and limitations. 
- Propose directions for future investigations. 
- This is a good place to re-state your findings as a final conclusion. What is the one take home message the reader should leave with?
- Your final conclusion needs to be very clear.

# References

- Provide a list of any references you use in the presentation. 